---
title: "Acoustic Correlates of Stress in Lithuanian"
subtitle: "Final Project Presentation"
author: "Chris Oakden"
date: "04/24/2018"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["http://www.jvcasillas.com/ru_xaringan/css/rutgers.css", "http://www.jvcasillas.com/ru_xaringan/css/rutgers-fonts.css"]
    nature:
      beforeInit: "http://www.jvcasillas.com/ru_xaringan/js/ru_xaringan.js"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```
```{r data/libraries, include = FALSE, warning=FALSE}
library(tidyverse)
data <- read_csv("../data_raw/M1Dur.csv")
data <- data %>%
  mutate(., f0_midpoint = as.double(f0_midpoint))
data2 <- read_csv("../data_raw/M1Int.csv")
```


#1. Introduction

--

###Acoustic Study of Stress/Tone in Lithuanian

--

-Production study

--

-Examine acoustic correlates of stress

--

-Increased **duration**

-Greater **intensity**

-Vowel quality (**F1/F2**) changes

-Higher pitch (**f0**)

--

###How is stress realized in a tone/pitch accent language?

---

#2. Preliminary Steps

--

###Data collection

--

-Record speakers in the Phonolab

--

-Segmentation of recordings using Praat TextGrids

--

-Extract acoustic values using Praat scripts

--

###Data organization

--

-Organize data, perform transformations, add additional variables in R

---

#3. Normalize duration within R

--

###Duration of fixed point in each frame sentence

--

-Find the **mean** of all fixed points

-Create a **multiplier** using the mean and each fixed duration

-Multiply that value by the duration of the vowels in the **corresponding** file

--

```{r fixes}
fix_1 <- data %>%
  filter(., TextGridLabel == "fix1") %>% 
  summarize(., mean = mean(duration)) %>%
  pull(.)

fix_2 <- data %>%                                                             
  filter(., TextGridLabel == "fix2") %>% 
  summarize(., mean = mean(duration)) %>%
  pull(.)
```

---

#3. Normalize duration within R

--

###Create two multiplier columns (one for each fixed point)

```{r fixes2}
data_mults <- data %>%   
  mutate(., multiplier1 = if_else(.$TextGridLabel=="fix1",                    
                                  fix_1 / duration, as.double(NA))) %>%        #create mult.
  fill(., multiplier1) %>%                                                     #fill in column
  mutate(., multiplier2 = if_else(.$TextGridLabel=="fix2",                     #create mult.
                                  fix_2 / duration, as.double(NA))) %>%             
  fill(., multiplier2, .direction = "up") %>%                                  #fill in column
  filter(., TextGridLabel != "fix1", TextGridLabel != "fix2")                  #delete fixed pts
```

---

#3. Normalize duration within R

--

###Multiply by the correct multiplier 

--

-using `case_when` and `stringr::str_detect`

```{r fixes3}
data_normdur <- data_mults %>%
  mutate(., NormalizedDuration = case_when(                                    #use case_when
    str_detect(.$TextGridLabel, "1$") ~ duration * multiplier1,                #use str_detect
    str_detect(.$TextGridLabel, "10$") ~ duration * multiplier1,               #and the last
    str_detect(.$TextGridLabel, "2$") ~ duration * multiplier1,                #strings of the
    str_detect(.$TextGridLabel, "20$") ~ duration * multiplier1,               #TextGrid labels
    str_detect(.$TextGridLabel, "3$") ~ duration * multiplier1,
    str_detect(.$TextGridLabel, "30$") ~ duration * multiplier1,
    str_detect(.$TextGridLabel, "00$") ~ duration * multiplier2,
    str_detect(.$TextGridLabel, "4$") ~ duration * multiplier2,
    str_detect(.$TextGridLabel, "40$") ~ duration * multiplier2,
    str_detect(.$TextGridLabel, "5$") ~ duration * multiplier2,
    str_detect(.$TextGridLabel, "50$") ~ duration * multiplier2,
    str_detect(.$TextGridLabel, "6$") ~ duration * multiplier2,
    str_detect(.$TextGridLabel, "60$") ~ duration * multiplier2,
    str_detect(.$TextGridLabel, "0$") ~ duration * multiplier1
  )) %>%
  select(., Filename, TextGridLabel, NormalizedDuration,                       #deselect 
         F1_midpoint, F2_midpoint, f0_midpoint)                                #multiplier cols
```

---

#4. Add new variables

--

###Add new variables for Vowel and Predicted Stress

```{r vars}
data_vars <- data_normdur %>%
  mutate(., Vowel = case_when(                                                 #Vowel
    str_detect(.$TextGridLabel, "^a") ~ "a",
    str_detect(.$TextGridLabel, "^e") ~ "e",
    str_detect(.$TextGridLabel, "^E") ~ "e",
    str_detect(.$TextGridLabel, "^i") ~ "i",
    str_detect(.$TextGridLabel, "^I") ~ "i",
    str_detect(.$TextGridLabel, "^o") ~ "o",
    str_detect(.$TextGridLabel, "^u") ~ "u",
    str_detect(.$TextGridLabel, "^U") ~ "u",
    str_detect(.$TextGridLabel, "^y") ~ "y"
  )) %>%
  mutate(., Stress = case_when(                                                #Predicted Stress
    str_detect(.$TextGridLabel, "^a") ~ "N",
    str_detect(.$TextGridLabel, "^e") ~ "N",
    str_detect(.$TextGridLabel, "^E") ~ "Y",
    str_detect(.$TextGridLabel, "^i") ~ "N",
    str_detect(.$TextGridLabel, "^I") ~ "Y",
    str_detect(.$TextGridLabel, "^o") ~ "N",
    str_detect(.$TextGridLabel, "^u") ~ "N",
    str_detect(.$TextGridLabel, "^U") ~ "Y",
    str_detect(.$TextGridLabel, "^y") ~ "N"
  ))
```

---

#4. Add new variables

--

###...and for repetition

--


```{r vars2}
data_rep <- data_vars %>%
    mutate(., Repetition = case_when(
    str_detect(.$Filename, "^M1A") ~ "1",
    str_detect(.$Filename, "^M1B") ~ "2",
    str_detect(.$Filename, "^M1C") ~ "3"
  ))
```


---

#5. Compile multiple dataframes

--

###Combine existing dataframe with file containing intensity data


--


```{r int}
data_intensity <- data2 %>%                                          
  filter(., TextGridLabel != "fix1", TextGridLabel != "fix2") %>%      #filter out fixed points
  select(., Intensity)                                                 #select intensity

data_final <- bind_cols(data_rep, data_intensity)                      #bind dataframes together
head(data_final)
```

---

#6. Preliminary EDA

--

###Basic generalizations from data ([u])

```{r udata, echo=FALSE}
udata <- data_final %>%
  filter(., TextGridLabel %in% c("u2", "U2", "u5", "U5",
                                "u20", "U20", "u50", "U50"))%>%
  filter(., Filename != "M1A74", Filename != "M1B25")
udata %>%
  group_by(., Stress) %>%
  summarize(., dur = (mean(NormalizedDuration)) * 1000,
               int = mean(Intensity),
               F1 = mean(F1_midpoint),
               F2 = mean(F2_midpoint),
               F0 = mean(f0_midpoint)) %>% knitr::kable(., format = 'html', digits = 2)
```

--

-Increased **duration**, **intensity**, **pitch** between stressed and unstressed [u]

--

-No real change in **F1**, ~200 Hz change in **F2**


---
#7. Next steps

--

###Linear mixed effects models

--

###Five models for each of two vowels [u] and [i]

--

###Criterion: acoustic correlate

--

###Predictors: **predicted stress**, # of syllables, segmental environment

--

###Random effects: # of syllables, segmental environment, repetition

--

###Isolate effect of predicted stress using **nested model comparison**

---
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
.center[
##Thank you!]


